---
title: "Automatically annotation contaminants"
author: "Jan Stanstrup"
date: "22 jul 2016"
output:
          html_document:
                        highlight: haddock
                        toc: yes
                        number_sections: yes
                        css: style.css
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Automatically annotation contaminants}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", dev='svg')
```



<style>
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all;
  word-wrap: initial; 
}
</style>



# Getting a list of known contaminants

We can pull a list of known contaminants from a list hosted on github (see `?get_cont_list`).
"Unknown" will used the positive mode list by default. This is mainly to do something when the ionization mode cannot be detected.


We need to load some libraries we will be using.

```{r libraries, message=FALSE, warning = FALSE}
library(MetabolomiQCsR)
library(purrr)
library(tidyr)
library(tibble)
library(dplyr)
library(magrittr)
library(knitr)
library(ggplot2)
library(plotly)
library(scales)
```


```{r contaminant list}
contaminant_list <- get_cont_list(polarity = c("positive", "negative", "unknown"))
contaminant_list %>% str
```


```{r contaminant list table}
contaminant_list[[1]] %>% 
                            slice(1:20) %>% 
                            select(-ESI, -MALDI, -References) %>% 
                            kable
```





# Read the raw files

Input files. These are local files for now.

```{r input files}
files <- c("Z:/_Data/LIP1/0000_Test/2016-06-24 - Contamination hunt - part I/Infusion tests/mzML_centroided/IPA_VWR_Boiled.mzML",
           "Z:/_Data/LIP1/0000_Test/2016-06-24 - Contamination hunt - part I/Infusion tests/mzML_centroided/EluentB_IPA1_Boiled.mzML",
           "Z:/_Data/LIP1/0036_SevaMeal/mzML/0036_LIP1p_20160201_034_2_1_1.mzML",
           "Z:/_Data/LIP1/0036_SevaMeal/mzML/0036_LIP1p_20160201_057_36_1_1.mzML"
)
```


```{r read raw files}
data_raw <- xcmsRaw_to_tbl(files)
```


# Get list of all known contaminants

Get list of contaminants for the polarities we have and add it to the file list.
```{r list of contaminants}
data_cont <-    data_raw %>% 
                distinct(polarity) %>% 
                mutate(cont_list = get_cont_list(polarity))
```




# Get EICs of all known contaminants

Clean up some column names and calculate EIC intervals.

```{r clean contaminant list}
EIC_ppm <- MetabolomiQCsR.env$target_cont$EIC_ppm # 30

data_cont %<>%  unnest(cont_list, .drop = FALSE) %>%  # one line per contaminant
                setNames(make.names(names(.))) %>% 
                mutate(comp_name = paste0(Compound.ID.or.species," (",Ion.type,", ", round(Monoisotopic.ion.mass..singly.charged.,4),")"), comp_mz = Monoisotopic.ion.mass..singly.charged. ) %>% 
                mutate(mz_lower = comp_mz-((EIC_ppm)/1E6)*comp_mz, mz_upper = comp_mz+((EIC_ppm)/1E6)*comp_mz) %>% 
                nest(-polarity, .key = contaminants)
```


Add the raw data to the table
```{r merge tables}
data <- left_join(data_raw,data_cont, by = "polarity")

rm(data_cont,data_raw)
```

Get EICs for all contaminants and raw files
```{r EIC for each contaminant}
data %<>%    mutate(   EIC = map2( raw, contaminants, get_EICs )   )
```


```{r EIC unnest contaminants and EIC}
data %<>%   mutate(EIC = map2(EIC,contaminants, ~ bind_cols(.y, data_frame(EIC = .x) ))) %>% 
            select(-raw,-contaminants) %>% 
            unnest

data %>% select(file, polarity, comp_name, comp_mz, mz_lower, mz_upper, EIC)  %>% print(width=Inf)
```



# Summarise contamination

```{r 'multiple  EIC summary'}
EIC_summary <-  data %>% 
                unnest %>% 
                group_by(file, comp_name, comp_mz, path) %>% 
                summarise(EIC_median = median(intensity), 
                          EIC_mean   = mean(intensity), 
                          EIC_sd     = sd(intensity), 
                          EIC_max    = max(intensity)
                          ) %>% 
                ungroup %>% 
                filter(EIC_median > 0)

EIC_summary %>% select(-path) %>% slice(1:10) %>% kable # kable is just to show a nice table below instead of the print display
```



```{r EIC plotting contamination profile, fig.height = 8, fig.width = 9}
EIC_summary %>% 
                mutate(sample = gsub(".mzML","",file)) %>% 
                mutate(sample = factor(sample,levels=unique(sample))) %>%
                select(sample,comp_name,EIC_median) %>%
                mutate(comp_name = factor(comp_name,levels=unique(comp_name)) ) %>%
                nest(-sample) %>% 
                mutate(gg = map2(data, sample, plot_contaminants))     ->
EIC_plots


EIC_plots %<>% mutate(  ggp = map(gg, plotly_build)  ) %>%
               mutate(  ggp = map(ggp, ~ plotly_clean_tt(.x, rep=c(`comp_name:` = "<b>Compound:</b>", `EIC_median:` = "<b>EIC Median:</b>"))    )  )

EIC_plots$ggp[[2]] 
```
